### 2017-08-03 analysis_dtree.tex

\section{Classification model} \label{classification_model}

In the section.~\ref{methodology} we discussed the basics of decision tree method and we provided a brief history of available algorithm. In this section we develop a classification model. Although there are numerous methods to classify labeled data, however, in this study we are interested in a method to not only predict the class of simulation based on metrics, but also give an intuition about the decision process. Therefor we use tree based algorithm to develop the prediction model.\\

After preprocessing data the most important step in developing a prediction model is feature selection. The step becomes more relevant when data set has hundreds to tens of thousands of variable or features. Although our dataset does not have numerous features we do some analysis on features before developing the prediction model. Features of a dataset could have similar behavior. This characteristic is easy to understand by correlation criteria. The Pearson correlation coefficient between two features is defined as

\begin{equation}
R(i)=\frac{cov(X_i,X_j)}{\sqrt{var(X_i)var(X_j)}},
\end{equation}

where $cov$ is covariance function and $var$ is the variance function \citep{Guyon_2003}. Fig.~\ref{fig:Corplot} represent the Pearson Correlation factor among features.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=200px]
        {figures/pdf/Figure_10.pdf}
    \caption{Pearson correlation values between metrics. The plot is generated using Corrplot package in R \citep{Corrplot_2016}}
    \label{fig:Corplot}
\end{figure}

As we can see from the figure some metrics are highly correlated. Among them we can mention C1 vs C2 , C3 vs C4, C5 vs C6 and C8, and C6 vs C8 have correlation factor more than 90\%. Also we observe that C11 is mostly have a negative correlation with other features. Feature selection is a broad research field. \citet{Guyon_2003} presented a review of common methods to select features, construct new features or reduce the domain. They highlighted the fact that redundant variable or variables that are useless by themselves may be useful in combination with other variables. Clustering is a method to construct new features as an unsupervised learning or feature selection without having the labels in calculation. The idea is to replace a group of features mean with newly developed features. It helps to reduce the number of features and use all of the available information \citep{Duda_1973,Guyon_2003}. Since we don't have a long list of features and also we want to study all feature effects we leave all features to be used in the prediction model.

After analysis of each features the next step is subsetting data for training and testing process. A subset of data that is used for training and test should have about the equal distribution of target values. Equally distributed data helps training process to not biased to an special class also helps to have an accurate evaluation of the model during the test process. Imbalanced data may decrease the classifier performance. A dataset is imbalanced if the classification categories are not approximately equally represented. \citet{Branco_2015} provided a comprehensive review of different approaches toward imbalanced distribution of target variables. A mostly common approach is data preprocessing through resampling. According to \citet{Branco_2015}, resampling strategies based on diverse set of techniques such as: random under/over-sampling, distance methods, data cleaning approaches, clustering algorithm among them.\\
According to the clustering results, among all data with different components and velocity model there are 816,1253,879, and 76 data belong to cluster 1,2,3, and 4, respectively. \citet{Weiss_2003} discussed the fact that the perfectly balanced dataset does not always provide the optimal results. One option could be under sampling cluster 1 to 3 to be in the similar distribution of cluster 4. This may eliminate the useful examples leading to a worse performance. Therefor we 10 times over-sample the data with cluster 4 to have a distribution similar to other cluster. Oversampling may increase the likelihood of overfitting, however, in this case the oversampling rate is not high. On the other hand, using pruning method during training as well as evaluating algorithm with unseen test dataset we address the overfitting issue.  Choosing what fraction of data should be used for training and testing is an open question. We put 30 percent of data for test and 70 percent of data for training.

Different measures are defined for evaluation of performance of prediction models. Two main metrics are precision and recall. In a simple word, in a binary case, precision represent the fact that how many of predicted values as positive is actually positive and recall represent the idea that how many of positive cases are diagnosticated with the algorithm. Choosing between these two measures is dependent on the application. For example if there is a highly contagious disease and we want to put all contaminated persons in a quarantine, we need to focus on recall value, because even if we miss one case, it could distribute the virus, however, there is a trade off between recall and precision. In the mentioned example there should be some person that indicated as contaminated with virus where as they are not. Therefor, the algorithm precision is not high. Table~\ref{tab:confusion_def} which is called confusion matrix is a method to represent the overall functionality of the algorithm \citep{Branco_2015}.

\begin{table}
\centering
\caption{Confusion Matrix for binary problem}
\label{my-label}
\begin{tabular}{llll}
\hline
                                                 &                              & \multicolumn{2}{c}{Prediction}                              \\ \cline{3-4} 
                                                 &                              & \multicolumn{1}{c}{Positive} & \multicolumn{1}{c}{Negative} \\ \cline{2-4} 
\multicolumn{1}{c}{Reference} & \multicolumn{1}{c}{Positive} & TP                           & FN                           \\
\multicolumn{1}{c}{}                             & \multicolumn{1}{c}{Negative} & FP                           & TN                           \\ \hline
\end{tabular}
\label{tab:confusion_def}
\end{table}

According to Table~\ref{tab:confusion_def} and the definition, mathematically precision and recall are defined as
\begin{equation}
Precision = \frac{TP}{TP+FP}, ~ ~ Recall = \frac{TP}{TP+FN},
\end{equation}

Because of trade of between these two metrics other metrics are defined to generate a metric considering both values. Among them we use $F-measure$ (Rijsbergen1979) which is defined as

\begin{equation}
F_\beta = \frac{(1+\beta)^2.recall.precision}{\beta^2.recall+precision},
\end{equation}

In this study we assume $\beta = 1$ to give same importance to precision and recall. \\
Developing a prediction model has a broad discussion. In this study we are not presenting the best model, however, we present an intuition into the method and be able to present a simple model where a user could estimate the target class without using sophisticated methods. We divided data into training and test set. C5.0 package in R has two options to prune the tree. These parameters include: Confidence factor (CF) and the smallest number of samples that must be put in at least two of the splits (minCases). We tune the algorithm for these two parameters based on training with train data and test the results on test data based on F1 score. We also active the winnow option, which activate the internal feature selection algorithm of the C5.0 package. Fig.~\ref{fig:tune} represent the tuning parameters results. It's obvious that the minCases is dominant factor. In this study we choose $CF=0.2$. 
\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_11.pdf}
    \caption{Results of tuning C5.0 algorithm through grid search on confidence factor (CF) and the smallest number of samples that must be put in at least tow of the splits (minCases). Color variation represents F1-score.}
    \label{fig:tune}
\end{figure}

Fig.~\ref{fig:tuneCFp2} represent the variation of F1-score for different clusters (or classes) with respect to the minCases.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=200px]
        {figures/pdf/Figure_12.pdf}
    \caption{Variation of F1-score for $CF=0.2$ for different cluster. Mean-F1 score represent the geometrical mean of F1-score of other clusters.}
    \label{fig:tuneCFp2}
\end{figure}

A decision tree could grow to very low levels and predict all training data accurately unless there is duplicated data with two different target value. The decision tree algorithm in this study provides very good results even the for test data. However, the scope of this study is to generate a simple relationship to evaluate the simulation results. In the case of going to very sophisticated methods, other algorithms (random forest, conditional random forest, neural networks, ...) could be much more accurate. Consequently, In this study we do not activate the boosting capability of the C5.0 algorithm. According to  Fig.~\ref{fig:tuneCFp2} lower minCases and higher CF provides a better results. It is worth to not that the results are based on evaluating algorithm on test data, therefore, higher values are not representing overfitting. However, as we mentioned earlier, we are not presenting the best model, rather we are presenting a simplest model that in general represent the feature effects. 
First we start with a very simple model. A simple model needs a strong pruning process.  Therefore we use $minCases = 100, CF=0.2$. Fig.~\ref{fig:dtree1} shows the classification decision tree. 
\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_13.pdf}
    \caption{First prediction model (M1) for classifying ground motion simulation process based on goodness of fit scores of two metrics (i.e., C8 (Response Spectra) and C4(Total Energy)). Number above the box represent the number of training data that end up to that specific node. Bar plots represent the probability of each class under the related condition.  }
    \label{fig:dtree1}
\end{figure}
Table~\ref{tab:confusionmat_test_1} shows  the confusion matrix of applying the prediction model on test dataset and Table~\ref{tab:prec_recall_test_1} represent the precision, recall, and F1-score for each individual class.

\begin{table}
\centering
\caption{Confusion Matrix for first prediction model.}
%\label{my-label}
\begin{tabular}{cccccc}
\hline
                           & \multicolumn{5}{c}{Prediction} \\ \hline
{Reference} & C  & 1    & 2    & 3    & 4    \\
                           & 1     & 240  & 25   & 0    & 0    \\
                           & 2     & 13   & 299  & 21   & 0    \\
                           & 3     & 0    & 38   & 211  & 9   \\
                           & 4     & 0    & 0    & 31  & 226  \\ \hline
\end{tabular}
\label{tab:confusionmat_test_1}
\end{table}

\begin{table}
\centering
\caption{Precision, Recall, and F1-score of the first prediction model calculated from the confusion matrix.}
\begin{tabular}{llll}
\hline
        & Precision & Recall  & F1 Score \\ \cline{2-4} 
Class 1 & 0.9056604 & 0.9486166 & 0.9266409 \\
Class 2 & 0.8978979 & 0.8259669 & 0.8604317 \\
Class 3 & 0.8178295 & 0.8022814 & 0.8099808 \\
Class 4 & 0.8793774 & 0.9617021 & 0.9186992 \\ \hline
\end{tabular}
\label{tab:prec_recall_test_1}
\end{table}

Summary of attribute usage in decision tree are according to Table~\ref{tab:attribute_usage_1}

\begin{table}
\centering
\caption{Percentage of data that used the attribute to classify them.}
\begin{tabular}{ccc}
\hline
Data percentage (\%) & Attribute & Metric                \\ \hline
100.00               & C8        & Response Spectra      \\
46.78                 & C4        & Total Energy         \\ \hline
\end{tabular}
\label{tab:attribute_usage_1}
\end{table}

Now we decrease the $minClass$ to 20 to have a more accurate model we call this model M2. Fig.~\ref{fig:dtree2} represents the resultant model.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_14.pdf}
    \caption{Second classification model}
    \label{fig:dtree2}
\end{figure}

Confusion matrix, model scores, and attribute usage are presented in  Table~\ref{tab:confusionmat_test_2},  Table~\ref{tab:prec_recall_test_2}, and  Table~\ref{tab:attribute_usage_2}.

\begin{table}
\centering
\caption{Confusion matrix}
%\label{my-label}
\begin{tabular}{cccccc}
\hline
                           & \multicolumn{5}{c}{Prediction} \\ \hline
{Reference} & C  & 1    & 2    & 3    & 4    \\
                           & 1     & 242  & 20   & 0    & 0    \\
                           & 2     & 11   & 330  & 15   & 0    \\
                           & 3     & 0    & 12   & 238  & 4   \\
                           & 4     & 0    & 0    & 10  & 231  \\ \hline
\end{tabular}
\label{tab:confusionmat_test_2}
\end{table}

\begin{table}
\centering
\caption{Recall and precision}
\begin{tabular}{llll}
\hline
        & Precision & Recall  & F1 Score \\ \cline{2-4} 
Class 1 & 0.9236641 & 0.9565217 & 0.9398058 \\
Class 2 & 0.9269663 & 0.9116022 & 0.9192201 \\
Class 3 & 0.9370079 & 0.9049430 & 0.9206963 \\
Class 4 & 0.9585062 & 0.9829787 & 0.9705882 \\ \hline
\end{tabular}
\label{tab:prec_recall_test_2}
\end{table}

\begin{table}
\centering
\caption{Data usage percentile}
\begin{tabular}{ccc}
\hline
Data percentage (\%) & Attribute & Metric                \\ \hline
100.00                 & C8        & Response Spectra      \\
100.00                 & C4        & Total Energy                \\ 
31.87                 & C6        & Peak Velocity      \\
25.66                 & C5        & Peak Acceleration      \\
19.50                 & C10        & Cross Correlation       \\\hline
\end{tabular}
\label{tab:attribute_usage_2}
\end{table}

% error and accuracy also are other measures.
%Rijsbergen, C. V. (1979). Information retrieval. dept. of computer science, university of glasgow, 2nd edition.