### 2017-09-29 acknowledgments.tex
completed abstract review, now working on introduction

\section{Acknowledgments}
\small

Acknowledgments here.