### 2018-02-05 method.tex
did some cleanng heare and there

\section{Data Analysis Method} 
\label{sec:method}

We are interested in proposing an algorithm with a reduced and prioritized number of validation metrics based on previously acquired validation data (i.e., our dataset). A common method to do this is to identify rules with disjunctive characteristics in the form of a decision tree that leads to an outcome representative of the overall analysis. In our case, we define such outcome in terms of four categories representative of the quality of the validation: poor, fair, good, excellent. The development of such a decision tree requires a proper classification of the data, which can be done through a clustering process. The following sections explain the data processing analysis we put in place to cluster the dataset and obtain two alternative decision trees.

\input{clustering}

\input{decision-tree}