### 2018-02-10 acknowledgments.tex
finished first complete version

\section{Acknowledgments}
\small

This work was supported by National Science Foundation (NSF) awards: ``SI2-SSI: Community Software for Extreme-Scale Computing in Earthquake System Science'' (ACI-1450451), and ``Improving Earthquake Forecasting and Seismic Hazard Analysis Through Extreme-Scale Simulations'' (OAC-1713792); and Southern California Earthquake Center (SCEC) awards: ``An Examination of the Correlations Between Different Goodness-of-Fit Metrics Based on a Large Dataset of Ground Motion Simulation Validations'' (16-067), and ``Application of Machine Learning in Deterministic Ground Motion Simulation'' (17-239). SCEC is funded by NSF Cooperative Agreement EAR-1600087 and U.S.~Geological Survey Cooperative Agreement G17AC00047. The SCEC contribution number for this paper is 8015.