### 2018-02-11 data-resources.tex

\section{Data and Resources}

Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed February 2018), and the C5.0 package developed by \citet{Kuhn_2017_Manual}. Additional calculations and figures were done using MATLAB (\url{http://www.mathworks.com}, last accessed February 2018). Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed February 2018). Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed February 2018). The validation dataset used here was readily available to the authors and can be provided without restrictions upon request.