### 2018-03-30 method.tex

\section{Data Analysis Method} 
\label{sec:method}

We are interested in proposing an algorithm with a reduced and prioritized number of validation metrics based on previously acquired validation data (i.e., our dataset). A common method to do this is to identify rules with disjunctive characteristics in the form of a decision tree that leads to an outcome representative of the overall analysis. In our case, we define such outcome in terms of four categories representative of the quality of the validation: poor, fair, good, excellent. The development of such a decision tree requires a proper classification of the data, which can be done through a clustering process. \myrevision{In this study we integrate three different topics (i.e., ground motion simulation, semi-supervised, and supervised learning), therefore, the same concept may have different technical terms in different part of the paper based on their application. In machine learning linguistic, attributes and features, mostly are used interchangeably. In this study, we use features for clustering and attributes for classification purposes. The reason for choosing these terms is keeping consistency with the references that we use to develop the methods. Therefore, in this study feature, attribute, and metric all refer to the eleven GOF scores. Consequently, dimension refers to the number of features in each constrained \kmeans{} clustering analysis. For example, 3-Dimensional subspace analysis is application of constrained \kmeans{} on three out of eleven features. Each observation with eleven metrics is categorized in one of Poor(p), Fair(F), Good(g), and Excellent (E) categories. The process of assigning these categories to each observation observation is called clustering or labeling. The resulted dataset is called the labeled dataset. When we use these labeled dataset in classification, the labels or clusters are called classes. Therefore, label, class, and cluster all refer to four categories of GOF level. }

The following sections explain the data processing analysis we put in place to cluster the dataset and obtain \myremove{two} \myrevision{three} alternative decision trees.

\input{clustering}

\input{decision-tree}