### 2018-04-27 results.tex

\section{Results}
\label{sec:results}

As explained in the methodology, the first step is to carry out a clustering process on the dataset in order to properly label the samples according to the classes to be used in the decision tree algorithm. Figure \ref{fig:clusters} presents a sample of the results obtained after applying the constrained \kmeans{} method, including subspacing. Recall that there are a total of 550 subspaces. Figure \ref{fig:clusters} shows three examples for each one of the subspaces considered, that is, three for each one of the 2-, 3-, and 4-feature subspace analyses. However, since it is not practical or possible to present three- or higher-dimensional plots, we display the results from a two-dimensional perspective into the subspaces by picking 2 metrics at a time. (Note that the only cases in which this is a direct view into the clustering results are the 2-feature analysis cases.)

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
	\caption{A sample of the results from the multi-dimensional clustering analysis showing the clusters from a two-dimensional perspective into the relationships between different GOF metrics. In each case, the labels near the upper-left corner indicate the features considered in the subspace analysis, and the features associated to the horizontal ($x$) and vertical ($y$) axes in the plot. The top, middle, and bottom rows correspond to 2-, 3-, and 4-feature subspace analyses. In each case, the poor, fair, good, and excellent clusters are indicated with circle, triangle, square, and cross symbols, respectively. The empty circles indicate the location of the artificial cannot-link stations we introduced as background knowledge to the clustering process. The color version of this figure is available only in the electronic edition.}
	\label{fig:clusters}
\end{figure*}

Some aspects of Fig.~\ref{fig:clusters} are worth discussing. Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case. In 
% the C5-C9 and C7-C9 combinations in 
\change{the combinations in}
the 3-feature analysis, for instance, the excellent cluster has a limited presence. This is due to the influence of the lower C9 values (see Fig.~\ref{fig:data-box-plot}) on the correlations with C5 and C7. This is not necessarily a bad thing, as it helps to understand the different relationships between the various GOF metrics. In the best of cases, the combinations C1-C2 and C5-C8 in the 2- and 4-feature analysis, for instance, exhibit an almost perfect proportionality between the metrics. That means either one of the metrics in those combinations is redundant. Relationships with redundant features are those where knowledge about one of the features provides a direct view into the other. On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant feature. We say C1 is irrelevant because it provides no insight about the outcome of C7 or that of the clusters. Identifying subspaces with redundant and irrelevant features is important because, on the one hand, the former help reduce the number of necessary features, while on the other hand, the latter can essentially be discarded because of their weak contribution to the decision making process \citep{Dy_2004_MLR}.

Figure \ref{fig:boxed-clusters} shows the statistical distribution (box-plots) of the samples once the dataset is partitioned into the four GOF validation classes. This is similar to Fig.~\ref{fig:data-box-plot}, but after the clustering process is completed. Separately, we prepared similar plots to look at the influence of the velocity models and components of motions on the results of the clustering process and, as observed before, they had no significant differences with the aggregate of all the samples in the dataset. \change{Figure \ref{fig:boxed-clusters} is particularly important because it highlights the outcome of the clustering process and provides a preview into the decision tree analysis results. Notice that metrics C5 through C8 consistently fall within the poor, fair, good, and excellent classification and that metrics C3 and C4 also show a progressive variation in sync with the validation classes. This means that these metrics are likely the best predictors for the outcome of the validation process, as we will see upon performing the decision tree analysis. On the other hand, metrics C1 and C2, and C9 through C11 are almost invariant, therefore less or not decisive in the validation process. C11, in particular, shows a broader (larger boxes) distribution, which indicates that it is less effective in predicting the final validation class.}

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-06}
	\caption{Statistical distribution of the dataset as partitioned into the four validation categories (poor, fair, good, excellent) after the clustering analysis. The distributions are shown in the form of box-plots for each metric (C1 through C11). In each case, \change{the medians are indicated by a notch in the boxes of the central quartiles, and the vertical} lines represent the interquartile range, with outliers shown as scattered dots. \change{The dashed lines and background shadows represent the boundaries of the poor, fair, good, and excellent categories as defined by \citet{Anderson_2004_Proc}.} The color version of this figure is available only in the electronic edition.}
	\label{fig:boxed-clusters}
\end{figure*}

In total, the clustering process results in 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively. These groups are shown in Fig.~\ref{fig:count-classes}. As it can be seen in this figure, the number of samples in the excellent class is significantly less than those in the poor, fair, and good classes. Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class\change{, for which we use the oversampling approach described in the previous section. Oversampling is nothing else but a replication of data samples. This process is done randomly, i.e., through a random selection of data samples from the original set to be duplicated until one increases the number of total samples in the oversampled set to a desired target number. In the case of the excellent class, we applied oversampling until reaching a total of 760 samples, as indicated with the dashed line in Fig.~\ref{fig:count-classes}. Since the original size was 76 samples, that means we applied an oversampling ratio of 10$\times$. According to \citet{Weiss_2003_JAIR}, an oversample rate of 10$\times$ is considered to be acceptable in this type of data processing.} (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary. Instead, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.

% We used the oversampling approach described in the previous section. We replicated data from the excellent class 
% \myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line. This oversampling is considered as a $\times$10 factor.} 
% \myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} 
% By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.45\textwidth]{figures/pdf/figure-07}
	\caption{Number of data samples in each class (poor, fair, good, excellent) after conducting a multi-dimensional constrained \kmeans{} clustering process using subspace analysis for 2, 3, and 4 features (GOF metrics). The dashed-line bar indicates the number of samples in the excellent class after oversampling. The color version of this figure is available only in the electronic edition.}
	\label{fig:count-classes}
\end{figure}

In total we generated 20,000 trees using the C5.0 algorithm for all possible combinations of the parameters CF and $S_{\min}$, where CF was chosen to vary between 0 and 1 at intervals of size 0.01, and $S_{\min}$ was chosen to vary between 1 and 200 at unitary intervals. Despite our choice for small intervals, the algorithm often reached recurrent tree structures for different CF and $S_{\min}$ values. Therefore, in reality, from the 20,000 combinations for which we ran the algorithm, only 66 unique trees were found.

For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth. Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of \change{decision} attributes \change{or GOF metrics} used in the trees as a function of the number of nodes and the depth of each tree. Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics). In general, all the trees obtained with the C5.0 algorithm are good in terms of the effectiveness factor ($F_1$ values close to 1). Then our choice comes down to using a reduced number of metrics. Having several trees with 2, 3, and 4 metrics (as opposed to 11), the following factor in the decision is choosing trees with algorithms using a small number of steps to reach the prediction. This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees. The rings around some of the dots indicate select trees used as reference in the text, other figures and tables. The color version of this figure is available only in the electronic edition.}
	\label{fig:nodes-depth}
\end{figure*}

\begin{table}[t]
	\centering
	\caption{Confusion matrix results for decision tree T1}
	\label{tab:t1:confusion}
	\small
	\begin{tabular}{cccccc}
		& 		& \multicolumn{4}{c}{Prediction}\\ 
		\cline{3-6}
		&   	& P  	& F     & G  	& E 	\\
		\cline{2-6}
		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
		& P 	& 242 	& 11 	& 0     & 0    	\\
		& F 	& 13 	& 313   & 16    & 0    	\\
		& G 	& 0 	& 42    & 207   & 15   	\\
		& E 	& 0 	& 0     & 23    & 231	\\ 
		\cline{2-6}
	\end{tabular}
\end{table}

\begin{table}[t]
	\centering
	\caption{Confusion matrix results for decision tree T3}
	\label{tab:t3:confusion}
	\small
	\begin{tabular}{cccccc}
		& 		& \multicolumn{4}{c}{Prediction}\\ 
		\cline{3-6}
		&   	& P  	& F     & G  	& E 	\\
		\cline{2-6}
		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
		& P 	& 248 	& 14 	& 0     & 0    	\\
		& F 	& 7 	& 343   & 18    & 0    	\\
		& G 	& 0 	& 9 	& 221   & 33   	\\
		& E 	& 0 	& 0     & 7    	& 213	\\ 
		\cline{2-6}
	\end{tabular}
\end{table}

Based on these criteria, we selected three candidate trees: T1, T2, and T3. These trees are shown in Fig.~\ref{fig:trees}. T1 is the simplest of the three, and T2 and T3 share some of their topology on the right-hand side. T3 is the most complex of them. More complex trees tend to be deeper, have more nodes, and employ more attributes\change{, all of which depends---in part---on the pruning process. More complex trees have} higher $F_1$ values, but they may not necessarily be more practical. In total T1, T2, and T3 use 2, 3 and 4 attributes respectively. All coincide in the use of the total energy (C4) and response spectra (C8) as key metrics. T2 adds peak acceleration (C5), and T3 adds peak velocity \change{(C6)} to the previous metrics.

At the bottom of each tree in Fig.~\ref{fig:trees} are the histograms of the data that land on each leaf node. (Note that the count of samples here is done based on the training dataset.) The decision tree assigns the final validation category based on the dominant class in each leaf. As such, the samples in the second leaf-node from the right in T3 are categorized as good despite a significant portion of them being excellent. This is a natural trade-off embedded in the use of decision trees.

% , which actually results in more effective trees.

\begin{figure*}
	\centering
	\includegraphics[width=0.86\textwidth]{figures/pdf/figure-09}
	\caption{Selected trees \textbf{(a)} T1, \textbf{(b)} T2, and \textbf{(c)} T3. In each case the decision nodes of the trees contain the code corresponding to the metric (see Table \ref{tab:metrics}) used and the branches beneath each node show the limit value to be used to select the next level in the tree. At the bottom, normalized histograms are shown to indicate the distribution of the samples at each leaf-node according to the validation classes with codes P, F, G, and E for poor, fair, good, and excellent, respectively. At the top of each histogram is the total count of samples at the corresponding leaf node. The histograms highlight the dominant validation class. The color version of this figure is available only in the electronic edition.}
	\label{fig:trees}
\end{figure*}

The actual effectiveness (i.e., the average $F_1$ values shown in Fig.~\ref{fig:nodes-depth}), however, is measured based on the testing dataset. Recall that $F_1$ depends on the number of true predictions and false predictions measured by the confusion matrix and the precision and recall factors. Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion} show the confusion matrices for T1 and T3, respectively; and Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} show the corresponding results for $P$, $R$, and $F_1$. As it can be seen in the Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion}, both trees lead to strong diagonal confusion matrices, meaning that the classification into poor, fair, good and excellent is well defined. The differences between both matrices are actually minor and small with respect to their diagonal values, and the $F_1$ values in Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} are all near to or above 0.9. The values obtained for tree T2 are consistent with \change{these observations}.

Another aspect of interest is the level of participation of each metric in the analysis of data, as the GOF values are run through \change{the nodes of the decision trees (i.e., as the tree is traversed)}. The results of such level of participation are listed in Table \ref{tab:participation} for trees T1, T2, T3, and T66. (The tree T66 is the most complex of them all as inferred from the number of metrics involved and the number of nodes and levels seen in Fig.~\ref{fig:nodes-depth}.) The percentages in this table represent the amount of data that is seen by a decision node associated to any particular metric. The nodes are not unique, and metrics are often used in different nodes in a given tree. Therefore, these percentage accumulate for each \change{metric as the tree is traversed}. A percentage of 100 means that a given metric has the opportunity to see all the data samples through one or multiple nodes, a low percentage means only a few data samples are evaluated by that metric, and a null percentage means the metric plays no role in the decision tree (it is not present in any node).

\begin{table}[t]
	\centering
	\caption{T1 precision, recall, and $F_1$ values per class}
	\label{tab:t1:f1}
	\small
	\begin{tabular}{lccc}
	Class   & $P$	& $R$  	& $F_1$ \\ 
	\hline
	P 		& 0.96 	& 0.95 	& 0.95 	\\
	F 		& 0.92 	& 0.86 	& 0.88 	\\
	G 		& 0.78 	& 0.84 	& 0.81 	\\
	E 		& 0.91 	& 0.94 	& 0.92 	\\
	\hline
	Mean 	&		&		& 0.90	\\
	\end{tabular}
\end{table}

\begin{table}[t]
	\centering
	\caption{T3 precision, recall, and $F_1$ values per class}
	\label{tab:t3:f1}
	\small
	\begin{tabular}{lccc}
	Class   & $P$	& $R$  	& $F_1$ \\ 
	\hline
	P 		& 0.95 	& 0.97 	& 0.95 	\\
	F 		& 0.93 	& 0.93 	& 0.93 	\\
	G 		& 0.84 	& 0.89 	& 0.86 	\\
	E 		& 0.96 	& 0.86 	& 0.91 	\\
	\hline
	Mean 	&		&		& 0.91	\\
	\end{tabular}
\end{table}

\change{The results in Table 6 highlight the fact that, for all trees,} the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms. They are followed by C5 (peak acceleration) and C6 (peak velocity). On the other end, C11 (strong phase duration) plays no role whatsoever in any of the trees, and that C1 (Arias integral), C2 (energy integral), and C3 (Arias intensity) have only small to marginal participations. The remaining metrics---C7 (peak displacement), C9 (Fourier spectrum), and C10 (Cross correlation)---have a somewhat significant participation, but only in the more complex trees.

\change{An additional aspect worth highlighting here is the fact that the clustering and decision tree processes address the problem of correlation between the metrics in a natural way. For instance, one would expect C3 and C5, or C3 and C4 to be related. Considering them all without distinction could lead to double weighing their influence, and assuming a 1-to-1 relationship may overstate their level of correlation. Instead, the process used here determines which of them has a better chance to predict the outcome in the presence of the other, and preserves that metric as part of the decision tree.}

\change{One can put this to test by removing one or more metrics and see how the participation of the metrics rearranges. Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing C8, and (ii) when removing C5 and C8. In the first case, C5 takes the place of the most determinant metric, followed by C4 and C6, depending on the tree. In the second case, C6 becomes the most dominant, followed by C3 and C4, which vary in their level of participation depending on the tree. These additional alternatives emphasizes the C4 and C6, and show that other metrics such as C5, C7 and C9 are also relevant. They also reaffirm the low or null levels of contribution of metrics C1, C2, C10 and C11. All this is consistent with what we observed in Fig.~\ref{fig:boxed-clusters}.}

% \myrevision{It is obvious that C8(response spectra) and C5(peak acceleration) are the most important metrics in developing the prediction models. Therefore, in order to understand other parameters importance we developed two sets of trees according to complexity of the original trees (the structure of generated trees are not presented here), while once dropping C8 (response spectra) and once dropping C8 (response spectra) and C5(peak acceleration) from the prediction attributes. According to Table \ref{tab:participation}, these trees also emphasize the importance of C4 and C6. However, in the absence of C5(peak acceleration) and C8 (response spectra),  C3 (Arias intensity) and C9 (Fourier spectrum) become more effective in developing prediction models. Also C11(strong phase duration) becomes somewhat important in the most complicated tree.}

\change{That being said, a} low participation does not necessarily mean that a given metric is not relevant at all. Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering. What the results we present here mean \change{is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more influential in the final result than the duration (C11). This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.}

% Original:
% A low participation does not necessarily mean that a given metric is not important. Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering. What the results we present here mean, in the context of GOF validation, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in reaching an accurate conclusion about validation quality than the duration (C11).

% Naeem suggestion:
% A low participation does not necessarily mean that a given metric is not important. Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering. What the results we present here mean, in the context of GOF validation for engineering applications, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in narrowing eleven validation metrics into fewer numbers in order to get the same categorization as all eleven metrics than the duration (C11).

In the end, the final selection of a preferred tree comes down to reducing complexity in the analysis without compromising the interpretation of results. We favor tree T1 because it is based only on three decision steps, and two GOF metrics: C4 (total energy), and C8 (response spectra).

\input{tab-partition}

\input{tab-partition-extra}