### 2018-05-23 data-resources.tex
changes before final submission

\section{Data and Resources}

Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed May 2018), and the C5.0 package (\url{https://CRAN.R-project.org/package=C50}, last accessed May 2018). Additional calculations and figures were done using MATLAB (\url{http://www.mathworks.com}, last accessed April 2018). Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed May 2018). Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed May 2018). The validation dataset used here was readily available to the authors and can be provided without restrictions upon request.